---
title: "Rowley - VWR - Phonological Priming"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r load packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(DT)
library(lme4)
library(scales)
library(broom)
library(sjPlot)
library(sjmisc)
```


# Deaf v. Hearing

## Load data 
We'll load the data from the CSV we created in `00_clean_data.R` and remove:

1. All rows with RTs < 250 or > 1200 using the `exclude_too_fast_slow` flag
2. All non-word trials 
3. Words: BADE, DUES, FOB, NORSE, VAT, WAIF
4. Subjects 9, 10, 32, 61

Below is a sample of that dataset: 

```{r load data, message=FALSE, warning=FALSE}
df <- read_csv("data/cleaned_dataset.csv") %>%
  filter(exclude_too_fast_slow == 0) %>%
  filter(word_type != 'Nonword') %>%
  filter(!(target_trial %in% c('BADE', 'DUES', 'FOB', 'NORSE', 'VAT', 'WAIF'))) %>%
  filter(!(subject %in% c(9, 10, 32, 61))) %>%
  mutate_at(vars(group, subject), factor) %>%
  rename(rt = reaction_time)

df %>%
  head() %>%
  datatable(options = list(dom = 't'))
```

## Subject-Level Data 

Let's take a look at our participants that we didn't remove:

```{r}
df %>%
  select(subject, age, group, handedness, sex) %>%
  distinct() %>%
  group_by(group) %>%
  summarise(participants = n(),
            age = round(mean(age), 1),
            male = sum(if_else(sex == 'male', 1, 0)),
            female = sum(if_else(sex == 'female', 1, 0)),
            right_handed = sum(if_else(handedness == 'right', 1, 0)),
            left_handed = sum(if_else(handedness == 'left', 1, 0))) %>%
  datatable(options = list(dom = 't'))
```

And their reaction time:

```{r}
# First calculate subject-level dataset 
subject_level <- df %>%
  group_by(group, subject, prime_type, word_type) %>%  # Subject-level
  summarise(accuracy = mean(accuracy),
            rt = mean(rt))

# Reaction time on group level
subject_level %>%
  group_by(group, prime_type, word_type) %>%  # Group-level
  summarise(rt = mean(rt)) %>%
  spread(word_type, rt) %>%
  ungroup() %>%
  mutate_if(is.numeric, round, 1) %>%
  datatable(options = list(dom = 't'))
```

And their accuracy: 

```{r}
# Accuracy on group level
subject_level %>%
  group_by(group, prime_type, word_type) %>%  # Group-level
  summarise(accuracy = mean(accuracy)) %>%
  spread(word_type, accuracy) %>%
  ungroup() %>%
  mutate_if(is.numeric, percent, .1) %>%
  datatable(options = list(dom = 't'))
```

## Visualization
```{r}
subject_level %>%
  ggplot(aes(x = group, y = rt, fill = prime_type)) +
  geom_boxplot() +
  scale_y_continuous()

subject_level %>%
  ggplot(aes(x = group, y = accuracy, fill = prime_type)) +
  geom_boxplot() +
  scale_y_continuous(labels = percent_format(1))


df %>%
  ggplot(aes(x = subject, y = rt, fill = group)) +
  geom_boxplot() +
  guides()
```

## Modeling 

To set up our linear mixed effects (also called hierarchical regression, multilevel models, etc.) we need to decide what are the fixed effects (the ones we are interested in) and what are the random effects (what we are not interested in, but which variance should be accounted for in our model).

Models for:

1. Reaction time
2. Accuracy

In each model, the fixed effects are:

1. Group (deaf/hearing)
2. Prime Type (phonological/graphemic)

And the random effects are:

1. List (A/B)
2. Subject
3. Item

### Reaction Time Model

```{r}
rt_model <- lmer(data = df, rt ~ group + prime_type + 
                                 (1|running_trial) + (1|subject) +
                                 (1|target_trial))

tab_model(rt_model)
plot_model(rt_model, vline.color = 'black',  show.values = TRUE)
```

### Accuracy Model

```{r}
acc_model <- lmer(data = df, accuracy ~ group + prime_type + 
                                 (1|running_trial) + (1|subject) +
                                 (1|target_trial))

tab_model(acc_model)
plot_model(acc_model, vline.color = 'black',  show.values = TRUE)
```

# 50 Hearing 

## Load data 
We'll load the data from the CSV we created in `00_clean_data_hearing.R` and remove:

1. All rows with RTs < 250 or > 1200 using the `exclude_too_fast_slow` flag
2. All non-word trials 
3. Subjects 2, 39, 126 for acc > |2.5| SD
4. Subject 1 (seems to be 2 participants under one ID)

**No words were excluded...that right?**

Below is a sample of that dataset: 

```{r}
hearing_df <- read_csv("data/cleaned_dataset_hearing.csv") %>%
  filter(exclude_too_fast_or_slow == 0) %>%
  filter(word_type != 'Nonword') %>%
  # filter(!(target_trial %in% c('BADE', 'DUES', 'FOB', 'NORSE', 'VAT', 'WAIF'))) %>%
  filter(!(subject %in% c(1, 2, 39, 126))) %>%
  mutate_at(vars(group, subject), factor) %>%
  rename(rt = target_displayed_rt_trial,
         accuracy = target_displayed_acc_trial)

df %>%
  head() %>%
  datatable(options = list(dom = 't'))
```


## Subject-Level Data

Let's take a look at our participants that we didn't exclude:

```{r}
hearing_df %>%
  select(subject, age, group, handedness, sex) %>%
  distinct() %>%
  summarise(participants = n(),
            age = round(mean(age), 1),
            male = sum(if_else(sex == 'male', 1, 0)),
            female = sum(if_else(sex == 'female', 1, 0)),
            right_handed = sum(if_else(handedness == 'right', 1, 0)),
            left_handed = sum(if_else(handedness == 'left', 1, 0))) %>%
  datatable(options = list(dom = 't'))
```

And their reaction time:

```{r}
# First calculate subject-level dataset 
subject_level <- hearing_df %>%
  group_by(subject, prime_type, word_type) %>%  # Subject-level
  summarise(accuracy = mean(accuracy),
            rt = mean(rt))

# Reaction time on group level
subject_level %>%
  group_by(prime_type, word_type) %>%
  summarise(rt = mean(rt)) %>%
  spread(word_type, rt) %>%
  ungroup() %>%
  mutate_if(is.numeric, round, 1) %>%
  datatable(options = list(dom = 't'))
```

And their accuracy: 

```{r}
# Accuracy on group level
subject_level %>%
  group_by(prime_type, word_type) %>%  # Group-level
  summarise(accuracy = mean(accuracy)) %>%
  spread(word_type, accuracy) %>%
  ungroup() %>%
  mutate_if(is.numeric, percent, .1) %>%
  datatable(options = list(dom = 't'))
```

## Visualization
```{r}
subject_level %>%
  ggplot(aes(x = prime_type, y = rt, fill = prime_type)) +
  geom_boxplot() +
  scale_y_continuous()

subject_level %>%
  ggplot(aes(x = prime_type, y = accuracy, fill = prime_type)) +
  geom_boxplot() +
  scale_y_continuous(labels = percent_format(1))

hearing_df %>%
  ggplot(aes(x = subject, y = rt, fill = group)) +
  geom_boxplot() +
  guides(fill = F) +
  theme(axis.text.x = element_text(angle = 90))
```

## Modeling 

To set up our linear mixed effects (also called hierarchical regression, multilevel models, etc.) we need to decide what are the fixed effects (the ones we are interested in) and what are the random effects (what we are not interested in, but which variance should be accounted for in our model).

Models for:

1. Reaction time
2. Accuracy

In each model, the fixed effect is:

1. Prime Type (phonological/graphemic)

And the random effects are:

1. List (A/B)
2. Subject
3. Item

### Reaction Time Model

```{r}
rt_model <- lmer(data = hearing_df, rt ~ prime_type + 
                                 (1|running_trial) + (1|subject) +
                                 (1|target_trial))

tab_model(rt_model)
plot_model(rt_model, vline.color = 'black',  show.values = TRUE)
```

### Accuracy Model

```{r}
acc_model <- lmer(data = hearing_df, accuracy ~ prime_type + 
                                 (1|running_trial) + (1|subject) +
                                 (1|target_trial))

tab_model(acc_model)
plot_model(acc_model, vline.color = 'black',  show.values = TRUE)
```

